---
title: "Machine Learning final assignment"
author: "AnwarSadique"
date: "April 14, 2016"
output: html_document
---
#Downloading and loading the files. Please make sure to move #the data to current working folder after downloading... The #below code to read.csv will not work if the files are not in #current working directory.
```{r}
#download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", method = 'curl')
#download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", method = 'curl')
rawDataTrain <- read.csv("pml-training.csv")
rawDataTest <- read.csv("pml-testing.csv")
dim(rawDataTrain); dim(rawDataTest)
```

#The amount of predictors are 160 which is too much. In an #effort to bring the numbers to down we will look at #predictors with lots of missing values and predictors with #very low variability. First i am checking predictors with low #variability using caret package nearzerovar function.
```{r}
library(caret)
rawDataTrain1 <- rawDataTrain[,-nearZeroVar(rawDataTrain)]
```

#Next I am going to check for missing values. I added missing #values of all remaining columns using sapply funciton, which #is 100 columns after eliminating the low variance columns.
```{r}
countNA <- sapply(rawDataTrain1, function(x) sum(is.na(x)) )
table(countNA)
```
#It is clear from the table function that there are 59 columns #with 0 NA values and 41 columns with 19216 NA values. 19216 #is approximately 97% of the total rows. All these columns can #be eliminated because 97% of all the value in those columns #are NA's.

```{r}
rawDataTrain2 <- rawDataTrain1[,countNA==0]
names(rawDataTrain2)
```

#Looking at the remaining columns it is clear that columns #with index which is column1 and columns 2,3,4,5 which has #date value of the data collected are not features. These are #information they might have entered when they collected the #dataset. Those columns are going to be eliminated from the #model building.
```{r}
training <- rawDataTrain2[,-c(1,2,3,4,5)]
remove("rawDataTrain1"); remove("rawDataTrain2");

```

#Doing the same for test data for the 20 questions to answer at the end of this assignment.
```{r}
rawDataTest1 <- rawDataTest[,-nearZeroVar(rawDataTrain)]
rawDataTest2 <- rawDataTest1[,countNA==0]
testData <- rawDataTest2[,-c(1,2,3,4,5)]
remove("rawDataTest1"); remove("rawDataTest2")

```

#I am going to use random forest to build the model. I will #partition the training and test data using 70% limit. I am #also going to do principle component analysis to reduce the #number of factors as 54 factors are way too much. 

```{r}
library(randomForest)
inTrain <- createDataPartition(training$classe, p=0.7, list=FALSE)
finalTrain <- training[inTrain,]
finalTest <- training[-inTrain,]
prProc <- preProcess(finalTrain[,-54], method = "pca", thresh=0.95)
pcaTrain <- predict(prProc, finalTrain[,-54])
rf <- randomForest(finalTrain$classe~., data = pcaTrain, verbose=TRUE, ntree=500)
print(rf)

```
#Now I will use test data to predict the training model. First #preprocess using the pcamodel of training data

```{r}
pcatest <- predict(prProc, finalTest[,-54])
confusionMatrix(finalTest$classe, predict(rf, pcatest))

```
#Now the final part is to predict the 20 differnt cases. We #need to do all preprocessing we did to the training set to #the 20 question set.

```{r}

pcaTestData <- predict(prProc, testData[,-54])
testData$classe <- predict(rf, pcaTestData)
testData$classe

```



